Wireless smart monitoring energy system: a proof of concept

### 2021-03-29 paper/main.tex
missing comma
ultime modifiche

	\section{State of the Art}
	Smart plugs with power consumption detection are commonly used in the same household scenarios, but they implement different workflows and protocols. In fact, usually to realize these IoT devices an external platform is integrated with the main board (Arduino, Raspberry Pi, generic boards with EPS 8266) . In most cases, this platform is Blynk, which is a hardware-agnostic IoT platform and allows to configure our board in many different ways.
	The user could use the cloud infrastructure provided by the Blynk company, or, since the source code is open-source, download and install their own server. 
	This latter scenario is the preferred one since is free, and consist of the Blink server, the blink app and the smart devices, which uses Blynk API to perform the connections. A Blynk installation is generic-purpose, so you could install and control different types of devices and show the measurements using only the TCP protocols. Anyway, app user interface needs to be customized and the server needs a database, so the configuration requires much time. Instead, our system - as described in the following paragraphs - has a different goal and provides essential features to manage the smart plugs. The aim of our project is to create a functional IoT system which avoids blackouts and in the meantime has a low impact on the wireless network using a modified version of the UDP protocol.
	Although, Blynk implements some security feature which makes it more secure respect to our system.

	\section{General architecture}
	The general architecture has three main components: the smart plugs, the server, and the web interface. The smart plugs code have been written using Arduino IDE in C++. Instead, the server has been developed in Java and the web interface is written in React. The rundown of what each part does is made up of a WeMos D1 R2, an Arduino-like board, monitors a sensor that gathers data on the power consumption, and controls a relay that turns on or off the appliance. It communicates over wireless with the UDP server that keeps track of the plugs and their energy consumption. The web interface uses an API exposed by the server to let the user see and control the status of each plug.
	\begin{figure}[htbp]
		\centering
		\includegraphics[width=\linewidth]{assets/architecture_schema}
		\caption{General architecture of the system}
		\label{fig:architecture_schema}
	\end{figure}

@@

	The connection scheme is represented in Figure~\ref{fig:pcbschema}, and has a cost of about €15, so it is not that expensive. The smart plug can be seen in Figure~\ref{fig:img5348}.
	The WeMos D1 R2 board is equipped with a EPS 8266 wireless module operating at 2.4GHz. The connection settings are hardcoded in the software because storing it would require an SD module and an SD card. The board is programmed in C/C++ using Arduino IDE and uses the JSON format to serialize the data that will be sent, using a UDP library, to the server.

@@

	Another thread, once the board has recognized and registered the server, sends updates periodically if there is a variation greater than the 10\% of the energy consumption.

@@

	Then, the plug starts to send \textit{UPDATE messages} to inform the server about energy consumption variation. To do so, a thread keeps watch over the readings provided by the PZEM sensor which integrates an ADC  system  in one integrated circuit \cite{8612412}.

@@

	We also used a different board called NodeMcu V3 in order to simulate a plug to get a real traffic scenario with at least two plugs. Since, NodeMcu V3 is not connected to a PZEM module, the power consumption readings are random values.  

@@

		\includegraphics[width=0.7\linewidth]{assets/plugserverinteraction}

@@

	\subsection{Server}\label{SVR}
	The server has been developed in Java 8 and managed using Maven. Tomcat is implemented to expose the servlet API. The server periodically broadcasts \textit{INIT packets} to let the plugs know its address and it listens for connections and updates from the plugs. Furthermore, it waits for acknowledgments and, in case they are not received re-transmits the lost packets. Finally, it exposes servlets to let the web interface get information, about the plugs and send \textit{ON/OFF packets}. The periodic actions, namely, sending the \textit{INIT packet} and re-transmitting those that are lost, are implemented as extensions of \textit{Guava's AbstractScheduledService} which allows performing of these at constant intervals. Another thread waits and manages the plugs replies. When a plug responds to the \textit{INIT packet}, the server will register its information and estimate the max power consumption for the appliance connected to the plug. This value is simply based on average values for the given appliance type as found on Google. When a plug sends an \textit{UPDATE packet} the server will update its registered information, most important: the current energy consumption and the max energy consumption. Finally, the server waits for the acknowledgments of previously ON or OFF packets. A summary of the logic used by the server is shown in Figure~\ref{fig:server_logic}. ON/OFF packets contain a \verb|timestamp| field which is sent back by the plugs in the acknowledgments. The server has a list of packets waiting for acknowledgment, in this way when the server receives an ACK it can discard the packet with the given timestamp. If a configurable number of seconds elapses before this happens the packet will be sent again. Like the plugs, the server first defines the content of the packets it sends as JSON objects. All the packets sent by the server have a field: \verb|act| this field can have as values:
	\begin{itemize}
		\item \begin{spverbatim}"INIT"\end{spverbatim} when broadcasting the INIT packet;
		\item \begin{spverbatim}"ON"\end{spverbatim} when sending a plug the instruction to turn on;
		\item \begin{spverbatim}"OFF"\end{spverbatim} when sending a plug the instruction to turn off.
	\end{itemize}
	\begin{figure}
		\centering
		\includegraphics[width=\linewidth]{assets/server_logic}
		\caption{Overview of server's logic for handling packets received by the plugs}
		\label{fig:server_logic}
	\end{figure}
	The servlets simply allow the web interface to get the server information about the plugs (the type of appliance connected, the current energy consumption, and more) and to turn on or off any given plug. In this last case, a packet will be sent by the server and for it, and this latter will wait for an acknowledgment. The information about the plugs saved by the server is kept as a \textit{thread-safe hashmap}. Some of this information, such as the IP address of the plugs, the type of appliance connected, and the max power consumption is also saved in a \textit{XML file} to avoid losing it in case of a server malfunction or restart.

@@

	\section{Challenges}
	During the development, we found some issues outside the scope of this POC. Since the focus of the project was not on security we decided for ease of debugging to do not encrypt the packets exchanged between plugs and server. Another issue is that the configuration of the plug is hard-coded, for example, the Wi-Fi SSID and password are stored as a char array in our C++ code. Since the EPS 8266 could switch to access point mode which means generate a new wireless network, a configuration procedure could be implemented. A device - such as a smartphone - could connect to the plug wireless network and send the SSID and password required by our home wireless settings. That information should be stored in a persistent drive - like a micro SD card.
	\section{Security issues}
	As we already mentioned, another challenge to face is the security issue, the information sent from a plug to the server is completely clear, so there is no encryption implemented. 
	In fact, our aim is to create a smart system and for debugging purpose is more convenient to use un-encrypted packets.
	 The following paragraphs describes different attacks which we have implemented or could be performed quite easily.
	\subsection{Fake plugs}
	A fake plug could be easily added to our system sending fake readings. In our scenario this could be a problem, because could lead to a different behavior.
	If an attacker uses a fake plug could send a random value to the system avoiding the user to turn on the correct device, considering that the system sums the energy consumption to prevent blackouts. An identification mechanism, for example using a token validation procedure, could be implemented to recognize and guarantee the access to authorized devices. An example of fake plug is the NodeMcu V3, which is used for debugging purpose, but from a different point of view could be see a malicious device and be a problem for a real scenario.
	\subsection{Man in the middle}
	A man in the middle attack could proxy the UDP traffic and change packets contents - such as the measurements or ON/OFF commands. 
	An attacker could even change the source IP address field of some INIT packets, and so the plugs would connect to the attacker server.
	Some tools, such as Scapy which is a Python library, allow to capture, modify and forward packets. Our system would not be able to detect that the packet has been modified, 
	also because the UDP checksum is not enabled and so, the integrity of header and payload is not guarantee.

	\section{System results}
	We used Wireshark to monitor our network while running the system. Generally, TCP traffic worsens the performances of UDP in wireless applications. In particular, during the test, we had both TCP and UDP third-party traffic, in particular, we had an Apex Legends game (UDP) and normal web navigation (TCP). The main objective of the test was to check the efficacy of our version of reliable UDP, so the focus was on the reception of acknowledgments by the server. The packets are sent to the correct client, even in a scenario with different traffic types. The traffic generated by the system is not much as can be seen in Figure~\ref{fig:grafico}.

>>>>

	\section{State of the Art}
	Smart plugs with power consumption detection are commonly
	used in the same household scenarios, but they implement
	different workflows and protocols. Usually to
	realize these IoT devices an external platform is integrated
	with the main-board (Arduino, Raspberry Pi, generic boards
	with EPS 8266). In most cases, this platform is Blynk, which
	is a hardware-agnostic IoT platform and allows us to configure
	our board in many different ways. The user could use the
	cloud infrastructure provided by the Blynk company, or, since
	the source code is open-source, download and install their server. This latter scenario is the preferred one since is free
	and consists of the Blink server, the blink app, and the smart
	devices, which use Blynk API to perform the connections. A
	Blynk installation is generic-purpose, so you could install and
	control different types of devices and show the measurements
	using only the TCP protocols. Anyway, the app user interface
	needs to be customized and the server needs a database, so
	the configuration requires much time. Instead, our system -
	as described in the following paragraphs - has a different goal
	and provides essential features to manage the smart plugs. Our project aims to create a functional IoT system that
	avoids blackouts and in the meantime has a low impact on
	the wireless network using a modified version of the UDP
	protocol. Although, Blynk implements some security feature
	which makes it more secure respect to our system.

	\section{General architecture}
	The general architecture, represented in Figure~\ref{fig:architecture_schema}, has three main components: the
	smart plugs, the server, and the web interface. The smart
	plugs code has been written using Arduino IDE in C++.
	Instead, the server has been developed in Java and the web
	interface is written in React. The rundown of what each part
	does is made up of a WeMos D1 R2, an Arduino-like board,
	monitors a sensor that gathers data on the power consumption,
	and controls a relay that turns on or off the appliance. It
	communicates over wireless with the UDP server that keeps
	track of the plugs and their energy consumption. The web
	interface uses an API exposed by the server to let the user see
	and control the status of each plug.
	\begin{figure}[htbp]
		\centering
		\includegraphics[width=\linewidth]{assets/architecture_schema}
		\caption{General architecture of the system}
		\label{fig:architecture_schema}
	\end{figure}

@@

	The connection scheme is represented in Figure~\ref{fig:pcbschema} and has a cost of about €15, so it is not that expensive. The smart plug can be seen in Figure~\ref{fig:img5348}.
	The WeMos D1 R2 board is equipped with an EPS 8266 wireless module operating at 2.4GHz. The connection settings are hard-coded in the software because storing it would require an SD module and an SD card. The board is programmed in C/C++ using Arduino IDE and uses the JSON format to serialize the data that will be sent, using a UDP library, to the server.

@@

	Another thread, once the board has recognized and registered the server, sends updates periodically if there is a variation greater than 10\% of the energy consumption.

@@

	Then, the plug starts to send \textit{UPDATE messages} to inform the server about energy consumption variation. To do so, a thread keeps watching over the readings provided by the PZEM sensor which integrates an ADC  system  in one integrated circuit \cite{8612412}.

@@

	We also used a different board called NodeMcu V3 to simulate a plug to get a real traffic scenario with at least two plugs. Since NodeMcu V3 is not connected to a PZEM module, the power consumption readings are random values.  

@@

		\includegraphics[width=\linewidth]{assets/plugserverinteraction}

@@

	\subsection{Server}\label{SVR}
	The server has been developed in Java 8 and managed using Maven. Tomcat is implemented to expose the servlet API. The server periodically broadcasts \textit{INIT packets} to let the plugs know its address and it listens for connections and updates from the plugs. Furthermore, it waits for acknowledgments and, in case they are not received re-transmits the lost packets. Finally, it exposes servlets to let the web interface get information, about the plugs and send \textit{ON/OFF packets}. The periodic actions, namely, sending the \textit{INIT packet} and re-transmitting those that are lost, are implemented as extensions of \textit{Guava's AbstractScheduledService} which allows performing of these at constant intervals. Another thread waits and manages the plugs replies. When a plug responds to the \textit{INIT packet}, the server will register its information and estimate the max power consumption for the appliance connected to the plug. This value is simply based on average values for the given appliance type as found on Google. When a plug sends an \textit{UPDATE packet} the server will update its registered information, most important: the current energy consumption and the max energy consumption. Finally, the server waits for the acknowledgments of previously ON or OFF packets. A summary of the logic used by the server is shown in Figure~\ref{fig:server_logic}. ON/OFF packets contain a \verb|timestamp| field which is sent back by the plugs in the acknowledgments. The server has a list of packets waiting for acknowledgment, in this way when the server receives an ACK it can discard the packet with the given timestamp. If a configurable number of seconds elapses before this happens the packet will be sent again. Like the plugs, the server first defines the content of the packets it sends as JSON objects. All the packets sent by the server have a field: \verb|act| this field can have as values:
	\begin{itemize}
		\item \begin{spverbatim}"INIT"\end{spverbatim} when broadcasting the INIT packet;
		\item \begin{spverbatim}"ON"\end{spverbatim} when sending a plug the instruction to turn on;
		\item \begin{spverbatim}"OFF"\end{spverbatim} when sending a plug the instruction to turn off.
	\end{itemize}
	\begin{figure}
		\centering
		\includegraphics[width=\linewidth]{assets/server_logic}
		\caption{Overview of server's logic for handling packets received by the plugs}
		\label{fig:server_logic}
	\end{figure}
	The servlets simply allow the web interface to get the server information about the plugs (the type of appliance connected, the current energy consumption, and more) and to turn on or off any given plug. In this last case, a packet will be sent by the server and for it, and this latter will wait for an acknowledgment. The information about the plugs saved by the server is kept as a \textit{thread-safe hashmap}. Some of this information, such as the IP address of the plugs, the type of appliance connected, and the max power consumption is also saved in an \textit{XML file} to avoid losing it in case of a server malfunction or restart.

@@

	\section{Challenges}
	During the development, we found some issues outside the scope of this POC. Since the focus of the project was not on security we decided for ease of debugging to do not encrypt the packets exchanged between plugs and server. Another issue is that the configuration of the plug is hard-coded, for example, the Wi-Fi SSID and password are stored as a char array in our C++ code. Since the EPS 8266 could switch to access point mode which means generate a new wireless network, a configuration procedure could be implemented. A device - such as a smartphone - could connect to the plug wireless network and send the SSID and password required by our home wireless settings. That information should be stored in a persistent drive - like a micro SD card.
	\section{Security issues}
	As we already mentioned, another challenge to face is the security issue, the information sent from a plug to the server is completely clear, so there is no encryption implemented. 
	In fact, we aim to create a smart system and for debugging purpose is more convenient to use un-encrypted packets.
	The following paragraphs describe different attacks which we have implemented or could be performed quite easily.
	\subsection{Fake plugs}
	A fake plug could be easily added to our system sending fake readings. In our scenario, this could be a problem, because could lead to different behavior.
	If an attacker uses a fake plug could send a random value to the system avoiding the user to turn on the correct device, considering that the system sums the energy consumption to prevent blackouts. An authentication mechanism, for example using a token validation procedure, could be implemented to recognize and guarantee access to authorized devices. An example of a fake plug is the NodeMcu V3, which is used for debugging purpose, but from a different point of view could be seen as a malicious device and be a problem in a real scenario. We used the NodeMcu V3 board to perform this attack, and since it is a high-level threat an authentication procedure is required.
	\subsection{Man in the middle}
	A man in the middle attack could proxy the UDP traffic and change the content of packets - such as the measurements or ON/OFF commands. 
	An attacker could even change the source IP address field of some INIT packets, and so the plugs would connect to the attacker server.
	Some tools, such as Scapy which is a Python library, allow to capture, modify and forward packets. Our system would not be able to detect that the packet has been modified, 
	also because the UDP checksum is not enabled and so, the integrity of the header and payload is not guaranteed.
	\subsection{Creating interference using a magnet}
	This last threat consists of using a magnet and gets it closer to the coil. The coil generates an electromagnetic field to measure the power consumption and using the magnets the measurements may change due to interference. This kind of attack works if there is no housing to protect the socket, and therefore also the coil. By using it, the magnet will not be able to alter the measurements. This all depends on the size of the magnet used, since we tried with a very small one and, pushing it away a little, there were no effects on the measurement. In the end, this attack is an open-ended question and needs further investigation.  
	\section{System results}
	Generally, TCP traffic worsens the performances of UDP in wireless applications. To check that this is not the case with our system, we used Wireshark to monitor our network while it was running to evaluate the impact of our traffic. During the test, we had both TCP and UDP third-party traffic, in particular, we had an Apex Legends game (UDP) and normal web navigation (TCP). The main objective of the test was to check the efficacy of our version of reliable UDP, so the focus was on the reception of acknowledgments by the server. The packets are sent to the correct client, even in a scenario with different traffic types. The traffic generated by the system is not much as can be seen in Figure~\ref{fig:grafico}. While the test was successful, it only had two plugs, one of them simulated. To achieve more realistic results the simulated plug generates more traffic than a real plug
